Topological Equivariant Artist Model

### 2021-03-26 slides/main.tex
everything but animated slides

\begin{frame}{Best practices in visualization design}
    \begin{description}
        \item[Expressiveness] structure preserving mappings\cite{mackinlayAutomatingDesignGraphical1986}
        \pause
        \item[Naturalness] easier to understand when properties match\cite{norman_things_smart}
        \pause
       \item[Graphical Integrity] graphs show \textbf{only} the data\cite{tufteVisualDisplayQuantitative2001}
    \end{description}
\end{frame}

@@

\begin{frame}{Topological Equivariant Artist Model}
    \centering
    An artist $\mathscr{\vartist}$ is an equivariant map 
    \begin{equation*}\Huge
        \mathscr{\vartist}: \mathscr{\dtotal} \rightarrow \mathscr{\gtotal}
    \end{equation*}
    from data $\mathscr{\dtotal}$ space to graphic $\mathscr{\gtotal}$ space. 
\end{frame}

\subsection{Data Model}
\begin{frame}{Model data as a fiber bundle \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}}
    A fiber bundle is a tuple $(\dtotal,\,\dbase,\,\pi ,\,\dfiber)$ defined by the map $\pi$
    \begin{equation*}
        \label{eq:fiber_bundle}
        \begin{tikzcd}[ampersand replacement=\&]
            \dfiber \arrow[r, hook] \& \dtotal \arrow[r, "\pi"] \& \dbase
        \end{tikzcd}
    \end{equation*}
    \begin{description}
        \pause
        \item[total space \dtotal] dataset topology
        \pause
        \item[fiber space \dfiber] variable schema
        \pause
        \item[base space \dbase] continuity
    \end{description}
\end{frame}

@@

\begin{frame}{Structure of components of \dfiber}
    \begin{figure}[H]
        \centering
        \includegraphics[height=.3\textheight]{figures/math/fiber.png}
    \label{fig:data_fiber_example}
    \end{figure}
        \begin{equation*}
            \dfiber = \dfiber_0 \times \ldots \times \dfiber_i \times \ldots \times \dfiber_n
        \end{equation*}
        \pause
        Monoid actions $\monoid_{i}$ (e.g. rotation, partial ordering) define the structure on $\dfiber_i$
        \begin{equation*}
            \bullet: \monoid_i\times \dfiber_i \rightarrow \dfiber_i
        \end{equation*}
        They must be associative and have an identity element. 
\end{frame}

@@

\subsection{Artist}%formalize the structure
\begin{frame}{Visual bundle $(\vtotal,\,\dbase,\,\pi ,\,\vfiber)$}
\begin{equation*}
    \label{eq:artist_diagram}
    \begin{tikzcd}[ampersand replacement=\&]
        \dtotal^{\prime} \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] \& \vtotal \arrow[d, "\pi"] \& \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] \& \gtotal \arrow[ld, "\pi"] \\
                                              \& \dbase                  \& \gbase \arrow[l, "\vindex"']                                              \&                    
        \end{tikzcd}
\end{equation*}
\pause
\begin{equation*}
    \mathcal{A}:\mathcal{E}\rightarrow\mathcal{E}
\end{equation*}
\pause
The topological artist is a sheaf map
\begin{equation*}
    \vartist: \mathcal{O}(\dtotal) \rightarrow \mathcal{O}(\gtotal)
\end{equation*}
\end{frame}

@@

\begin{frame}{Stevens' Scales \cite{stevensTheoryScalesMeasurement1946}}
\begin{table}[H]
    \begin{tabularx}{\textwidth}{|l|l|X|}\toprule
        \textbf{scale} & \textbf{group} & \textbf{constraint} \\\midrule
        nominal & permutation &  $\text{if } \delement_1 \neq \delement_2 \text{ then } \vchannel (\delement_1) \neq\vchannel(\delement_2)$\\
        ordinal &  monotonic & $\text{if } \delement_1 \leq \delement_2 \text{ then } \vchannel (\delement_1) \leq \vchannel(\delement_2)$\\
        interval &  translation &  $\vchannel (x + c) = \vchannel(x) + c$ \\
        ratio &  scaling &  $\vchannel(xc) = \vchannel(x)*c $\\ \bottomrule
    \end{tabularx}
\end{table}
\end{frame}

\begin{frame}{Equivariance: Partial Order}
    \begin{figure} %break out steps/reveal
        \includegraphics[width=1\textwidth]{figures/math/monoid_equivariant.png}
    \end{figure}
\end{frame}

\begin{frame}{Visualization Equivariance}
    \begin{figure}[H]
        \includegraphics[width=\textwidth]{figures/math/diff_type_q.png}
    \end{figure}

    %put equation w/ rho, label rho, rho'
\end{frame}

\begin{frame}{Scatter: $\vmark(xpos, ypos)(\alpha, \beta)$}
    \begin{figure}[H]
        \includegraphics[width=1\textwidth]{figures/math/scatter.png}
    \end{figure}

    \begin{align*}
        x &= size *\alpha \cos(\beta) + xpos \\
        y &= size *\alpha \sin(\beta) + ypos
    \end{align*}    
\end{frame}

@@

\begin{frame}{Build \vmark over \dbase:\vmarkd}
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/q_hat.png}
\end{figure}
\begin{equation*}\label{eq:qhat_q_s}
    \vmarkd(\vsection(\dbasepoint))(\gbasepoint) \coloneqq \vmark((\vsectionpull)(\gbasepoint))
\end{equation*} 
such \gbasepoint\ can be factored out when $\vpreimg = \gbasepoint$
\end{frame}

\begin{frame}{Composition of artists} 
    % make disjointed figure or scatter + line
    Given the family of artists $(\dtotal_i: i\in I)$ on the same image
    \begin{equation*}
        + \coloneqq \underset{i\in I}{\sqcup} \dtotal_{i}
    \end{equation*}
    the + operator defines a simple composition of artists. 
\end{frame}

@@

\begin{frame}[fragile]{Artists: Scatter \& Line}
%%keep the signatures, comments, change names to math
\begin{minted}{python}
    class Point(mcollections.Collection):
        def assemble(self, x, y, s, facecolors='C0' ):
            # construct geometries of the circle glyphs in visual coordinates
            self._paths = [mpath.Path.circle(center=(xi,yi), radius=si) 
                        for (xi, yi, si) in zip(x, y, s)] 
            # set attributes of glyphs, these are vectorized 
            # circles and facecolors are lists of the same size
            self.set_facecolors(facecolors)

    class Line(mcollections.LineCollection):
        def assemble(self, x, y, color='C0'):
            # assemble line marks as set of segments 
            segments = [np.vstack((vx, vy)).T for vx, vy in zip(x, y)]
            self.set_segments(segments)
            self.set_color(color)
    \end{minted}
\end{frame}

@@

\begin{frame}{Proposed Dissertation}
% math + prototypes, organize it
% papers + contributions to matplotlib
\begin{itemize}
    \item expansion of the mathematical framework to include worked out simple and complex addition
    \item formalization of definition of equivalance class \vartisteq
    \item implementation of artist with explicit \vindex\
    \item specification of interactive visualization
    \item mathematical formulation of a graphic with axes labeling
    \item implementation of new prototype artists that do not inherit from Matplotlib artists
    \item provisional mathematics and implementation of user level composite artists
    \item proof of concept domain specific user facing library 
\end{itemize}
\end{frame}

>>>>

\begin{frame}{Best practices in visualization design}
    \begin{description}
        \item[Expressiveness] structure preserving mappings\cite{mackinlayAutomatingDesignGraphical1986}
        \pause
        \item[Graphical Integrity] graphs show \textbf{only} the data\cite{tufteVisualDisplayQuantitative2001}
        \pause 
        \item[Naturalness] easier to understand when properties match\cite{norman_things_smart}
    \end{description}
\end{frame}

@@

\begin{frame}{Topological Equivariant Artist Model}
    \centering
    An artist $\mathscr{\vartist}$ is an equivariant map 
    \begin{equation*}\huge
        \mathscr{\vartist}: \mathscr{\dtotal} \rightarrow \mathscr{\gtotal}
    \end{equation*}
    from data $\mathscr{\dtotal}$ space to graphic $\mathscr{\gtotal}$ space. 
\end{frame}

\subsection{Data Model}
\begin{frame}{Model data as a fiber bundle \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}}
    A fiber bundle is a tuple $(\dtotal,\,\dbase,\,\pi ,\,\dfiber)$ defined by the map $\pi$
    \begin{equation*}
        \label{eq:fiber_bundle}
        \begin{tikzcd}[ampersand replacement=\&]
            \dfiber \arrow[r, hook] \& \dtotal \arrow[r, "\pi"] \& \dbase
        \end{tikzcd}
    \end{equation*}
    \begin{description}
        \pause
        \item[total space \dtotal] topology
        \pause
        \item[fiber space \dfiber] schema
        \pause
        \item[base space \dbase] continuity
    \end{description}
\end{frame}

@@

\begin{frame}{Structure of components of \dfiber}
    \begin{figure}[H]
        \centering
        \includegraphics[height=.3\textheight]{figures/math/fiber.png}
    \label{fig:data_fiber_example}
    \end{figure}
        \begin{equation*}
            \dfiber = \dfiber_0 \times \ldots \times \dfiber_i \times \ldots \times \dfiber_n
        \end{equation*}
        \pause
        Monoid actions $\monoid_{i}$ (e.g. rotation, partial ordering) define the structure on $\dfiber_i$
        \begin{equation*}
            \bullet: \monoid_i\times \dfiber_i \rightarrow \dfiber_i
        \end{equation*}
        where $\bullet$ is associative and has an identity action. 
\end{frame}

@@

\subsection{Artist}
\begin{frame}{Visual bundle $(\vtotal,\,\dbase,\,\pi ,\,\vfiber)$}
    \begin{equation*}
        \mathcal{A}:\mathcal{E}\rightarrow\mathcal{H}
    \end{equation*}
\pause
\begin{equation*}
    \label{eq:artist_diagram}
    \begin{tikzcd}[ampersand replacement=\&]
        \dtotal^{\prime} \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] \& \vtotal \arrow[d, "\pi"] \& \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] \& \gtotal \arrow[ld, "\pi"] \\
                                              \& \dbase                  \& \gbase \arrow[l, "\vindex"']                                              \&                    
        \end{tikzcd}
\end{equation*}
\pause
\begin{equation*}
    \vartist: \mathcal{O}(\dtotal) \rightarrow \mathcal{O}(\gtotal)
\end{equation*}
\end{frame}

@@

\begin{frame}{Group Equivariance: Stevens' Scales \cite{stevensTheoryScalesMeasurement1946}}
\begin{table}[H]
    \begin{tabularx}{\textwidth}{|l|l|X|}\toprule
        \textbf{scale} & \textbf{group} & \textbf{constraint} \\\midrule
        nominal & permutation &  $\text{if } \delement_1 \neq \delement_2 \text{ then } \vchannel (\delement_1) \neq\vchannel(\delement_2)$\\
        ordinal &  monotonic & $\text{if } \delement_1 \leq \delement_2 \text{ then } \vchannel (\delement_1) \leq \vchannel(\delement_2)$\\
        interval &  translation &  $\vchannel (x + c) = \vchannel(x) + c$ \\
        ratio &  scaling &  $\vchannel(xc) = \vchannel(x)*c $\\ \bottomrule
    \end{tabularx}
\end{table}
\end{frame}

\begin{frame}{Monoid Equivariance: Partial Orders}
    \begin{figure} %break out steps/reveal
        \includegraphics[width=1\textwidth]{figures/math/monoid_equivariant.png}
    \end{figure}
\end{frame}

\begin{frame}{Visualization Equivariance}
    \begin{figure}[H]
        \includegraphics[width=\textwidth]{figures/math/diff_type_q.png}
    \end{figure}
    \begin{equation*}
        \vmark(\vsection) = \vmark(\vsection^{\prime})\implies \vmark(m\circ\vsection) = \vmark(m\circ\vsection^{\prime})
    \end{equation*}

\end{frame}

\begin{frame}{Scatter: $\vmark(xpos, ypos)(\alpha, \beta)$}
    \begin{figure}[H]
        \includegraphics[width=1\textwidth]{figures/math/scatter.png}
    \end{figure}   
\end{frame}

@@

\begin{frame}{Build \vmark\ over \dbase:\vmarkd}
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/q_hat.png}
\end{figure}
\begin{equation*}\label{eq:qhat_q_s}
    \vmarkd(\vsection(\dbasepoint))(\gbasepoint) \coloneqq \vmark((\vsectionpull)(\gbasepoint))
\end{equation*} 
\end{frame}

\begin{frame}{Composition of artists $+ \coloneqq \sqcup \dtotal_{i}$} 
    \begin{figure}
        \includegraphics[width=1\textwidth]{figures/math/exploding_artist.png}
    \end{figure}
\end{frame}

@@

\begin{frame}[fragile]{Artists: Scatter \& Line \vmarkd}
\begin{minted}{python}
    class Point(mcollections.Collection):
        def assemble(self, x, y, s, facecolors='C0' ):
            # construct geometries of the circle glyphs in visual coordinates
            # set attributes of glyphs

    class Line(mcollections.LineCollection):
        def assemble(self, x, y, color='C0'):
            # assemble line marks as set of segments 
    \end{minted}
\end{frame}

@@

\begin{frame}{Proposed Work}
\begin{figure}
    \includegraphics[width=1\textwidth]{figures/intro/gantt.png}
\end{figure}
\end{frame}
